### 2019-04-22 tex/recap.tex
Update recap.tex

\documentclass[dvipsnames,a4paper,11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{todonotes}
\usepackage{xcolor}
\usepackage{tabularx}

\newcommand{\tiph}[1]{\todo[inline,color=Orchid]{#1}}
\newcommand{\pimprenelle}[1]{\todo[inline, color=ForestGreen]{#1}}

\title{Multiplex stream graphs : tableau récapitulatif des notions existantes pour les streams et les multilayer et pour les Stream-multilayer}
\author{Pimprenelle Parmentier, Tiphaine Viard, ...}

\begin{document}
    \maketitle
	\begin{tabularx}{\linewidth}{|X||X|X||c|}
		\hline
		\textbf{Classical Graphs} & \textbf{Stream graphs} & \textbf{Multilayer} & \textbf{Multilayer stream graphs}\\
		$G=(V,E)$ & $S = (T,V,W,E)$ &  $M= (V_M,E_M,V,L)$ & $M = (T,T_M,V,W_M,E_M,{\cal L})$\\
		\hline
		\hline
		- & - & intra-layer graph & intra-layer graph\\
		\hline
		- & - & inter-layer graph & inter-layer graph \\
		\hline
		\hline
		- & covergage & - & intra/inter layer coverage \\
		- & cov$(S)=\frac{|W|}{|T\times V|}$&-& (importance de garder les $V_M$ ?)\\
		\hline
		- & $G_t$ graph at $t$ & - & $M_t$ \\
		\hline
		- & Graph induced & - & Multilayer induced \\
		\hline
		- & $G(S)$ & - & $M_I(M)$\\
		\hline
		- & - & underlying graph & underlying Stream \\
		- & - & $G(M)$ & $S_U$\\
		\hline
		Number of nodes & Number of nodes & Number of nodes & Number of nodes \\
		$|V|$ & $\frac{|W|}{|T|}$ & $|V|$ & $\frac{|W|}{|T|}$??? problem\\
		\hline
		number of links & number of links & number of links & number of links\\
		$|E|$ & $\frac{|E|}{|T|}$ & $|E_M|$ & ??? to do\\
		\hline
		- & compactness & - & compactness \\
		 & $\frac{|W|}{|T'\times V'|}$& &???\\
		 \hline
		 - & uniformity & - & uniformity \\
		 & $U = \frac{\sum_{uv \ V \otimes V}{|T_u \cap T_v|}}{\sum_{uv \in V\otimes V}{|T_u\cup T_v|}}$& & ??? to do\\
		 \hline
		 density & density & ?? & density ? \\
		 $\frac{|E|}{|V|\otimes V|} = \frac{|E|2}{|V|(|V|-1)} $& $\delta(S) = \frac{\sum_{uv \in V \otimes V}{|T_{uv}|}}{\sum_{uv \in V\otimes V}{|T_u\cap T_v|}}= \frac{\int_{t\in T}{|E_t|dt}}{\int_{t\in T}{|V_t\otimes V_t|dt}}$ & & \\
		 \hline
		 \hline

	\end{tabularx}
	- : doesn't exists
\end{document}